Robust Control of a Fluxonium Qubit

### 2020-10-12 paper/main.tex
updates

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity operations on quantum processors in the presence of
  calibration errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
  but they require exquisite calibration to be performant. In this work we employ robust trajectory optimization techniques
  to achieve high fidelity gates for parameter spreads on the order of $1\%$. 
  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
  the Lindblad master equation, which increases the computational complexity
  of the optimization algorithm.
  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
  for the gate set of a fluxonium qubit
  using a computationally efficient metric in optimization.
\end{abstract}

@@

%% S2
\section{QOC + AL-iLQR}
(QOC Problem Statement) Here we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schroedinger equation (TDSE)
\label{eq:tdse}
\begin{align}
  \frac{d}{dt} \ket{\psi(t)} &= -\frac{i}{\hbar} H(u(t), t) \ket{\psi(t)}
\end{align}
The evolution is sometimes cast with the evolution
of a density matrix under the Lindblad master equation to
model the decoherence of the state explicitly. The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system.

@@

(AL-iLQR Problem Statement) The trajectory optimization
literature solves a more general class of non-linear programs that resemble
the quantum optimal control problem. The quantum optimal control
problem is a specific case of the linear quadratic regulator (LQR).
LQR is so called because the dynamics are linear in the state and
the functional is quadratic in the state. In the LQR formulation
the same functional is evaluated at each knot point
\begin{equation}
  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
\end{equation}
where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
at knot point $k$ and final state, $u_{k}$ are the controls,
and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
The state is propagated using a dynamics function
$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
and $f$ encodes the TDSE dynamics. In the following
we refer to $\ket{\psi}$ as the state. We refer to $x$ and $u$ as
the augmented state and augmented controls, respectively.

@@

%% S4
\section{Longitudinal Relaxation Awareness}
(Strategy) We seek to minimize the probability
that the qubit decays as a result of longitudinal
relaxation. To this end we add the
longitudinal relaxation probability
to the augmented state vector
\begin{equation}
  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
\end{equation}
Setting the target
longitudinal relaxation probability to 0 results in
a quadratic cost at each knot point
of the form ${\lvert P_{1}(t_{k}) \rvert}^{2}$.

$T_{1}(a_{k})$
is obtained at each knot point by evaluating
a spline interpolant fit to
experimentally obtained data of the form
$\{(a, T_{1})\}$.
Calculating $T_{1}$ directly from theoretical
considerations requires many high-rank
eigendecompositions, which
is computationally expensive. Additionally,
$T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data increases
the fridge truth of the simulation.

For a given $T_{1}$ time, a shorter gate duration would tend to
favor a lower longitudinal relaxation probability.
We allow the optimizer to tune the gate duration by
making the time step between each knot point $\Delta t_{k}$
a decision variable.
Promoting $\Delta t_{k}$ to a decision variable, rather
than the number of knot points $N$, preserves the
Markovian decision structure of the trajectory
optimization problem. To ensure numerical
integration accuracy is maintained we add a bound
constraint at each knot point
$2.5\textrm{e-}3 \ \textrm{ns} \le
\Delta t_{k} \le 1\textrm{e-}2 \ \textrm{ns}$.
Note that this bound constraint is allowed to be
broken for intermediate iterations of the optimization,
so we add the square root of the time step $\sqrt{\Delta t_{k}}$
to the augmented control vector and use the squared root
of the time step $\lvert \Delta t_{k} \rvert$ in optimization.

(Results) The $Y/2$ and $Z/2$ gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$ (see Figure 1). Both extend their gate times
beyond their analytic counterpart, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. Both gates reduce
their probability of longitudinal relaxation per gate by a factor of $\sim 5$ over
their analytic counterparts. Furthermore, they perform simlarly in
the concatenated gate application comparision, maintaining a total gate error under $10^{-2}$
over $10^{3}$ gate applications, for an average gate error on the order of $10^{-5}$.
We define gate error as the infidelity of the final state and the target state,
averaged over multiple pseudo-randomly
generated initial states. The $X/2$ gate

%% T1
%% TODO: maybe for appendix.
\begin{table}[ht]
  \begin{tabular}{c | c | c | c}
    & Analytic  & QOC &\\
    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$\\
    \hline
    Z/2 & 5.745 & 1.149 & 5.000\\
    Y/2 & 5.253 & 1.157 & 4.540\\
    X/2 & 16.251 & 4.488 & 3.621\\
  \end{tabular}
  \caption{Probability of longitudinal relaxation for each gate
    evaluated at the gate's duration.}
\end{table}

@@

  \caption{
    (a) Z/2 gates constructed using the
    analytic, 2-point sampling, 8-point unscented sampling, and the 1\textsuperscript{st}
    and 2\textsuperscript{nd} order derivative methods. The gates
    shown for the robust methods are the solutions at twice the analytic
    gate time.
    (b) Single gate error as
    a function of the gate duration at a one-percent
    detuning from the nominal qubit frequency for all methods.
    (c) Single gate error as a function of the detuning from the nominal
    qubit frequency at multiples of the analytic gate duration for the analytic
    and 1\textsuperscript{st} order derivative methods.
  }
\end{figure*}

@@

%% S5
\section{Robustness to Static Parameter Deviations}
(Strategy) Uncertainty in the Hamiltonian parameters lead
to erroneous 
The uncertainty Hamiltonian parameters are known to
some finite experimental accuracy. Further,
experimental fluctuations may cause parameters to deviate
from their
the qubit frequency to deviate from its measured value
by an amount $\delta f_{q}$. There are other situations that
arise where this is useful, such as static amplitude
offsets or static gain offsets.

System parameter
deviations are typically addressed with dynamic decoupling
sequences \cite{merrill2014progress},
the DRAG scheme \cite{krantz2019quantum}, or
geometric phase considerations
\cite{xu2020nonadiabatic} \cite{han2020experimental}.
Dynamic decoupling sequences compose rotations on the
Bloch sphere so that erroneous rotations arising due
to the parameter deviation are cancelled.
%% The latter two methods are inflexible.
We draw on the
robust control literature to demonstrate
numerical techniques for engineering robustness
to parameter deviations.

We propose two methods for engineering robustness
to parameter deviations. The first is the
derivative method. We draw on the intuition that
the sensitivity of the state evolution to the parameter
$\lambda$ is encoded in the derivative of
the state with respect to that parameter
$\partial_{\lambda}^{l} \ket{\psi}$. The derivative
scheme aims to minimize the norm of this derivative.
The derviatives of the state with respect to the parameter
of interest are propgated in the
augmented state vector. Their dynamics are found by differentiating
the TDSE dynamics with respect to the parameter of interest
(see Appendix). Setting the target derivatives of the state
to the zero vector results in quadratic costs at each
knot point
${\lvert \partial_{\lambda}^{l} \ket{\psi} \rvert}^{2}$.
In the $l$\textsuperscript{th}-order derivative method, the norm of
all derivatives of order less than or equal to $l$ are penalized. For example,
in the 2\textsuperscript{nd}-order derivative method both the
1\textsuperscript{st}-order state derivative
(${\lvert \partial_{\lambda} \ket{\psi_{k}} \rvert}^{2}$)
and the 2\textsuperscript{nd}-order state derivative
(${\lvert \partial^{2}_{\lambda} \ket{\psi_{k}} \rvert}^{2}$)
contribute quadratic costs at each knot point.

The second scheme we analyze is the sampling method,
which is well studied in the robust control literature
\cite{manchester2016derivative} \cite{tronarp2016sigma}.
We add additional
states to the augmented state vector that follow deviant dynamics, and
penalize the state's difference from the target state.
The deviant dynamics are given by the system
Hamiltonian with the parameter of interest replaced by
a deviant value $\lambda \gets \lambda + \delta \lambda$
The parameter's deviation is typically chosen
to represent the statistical distribution of the parameter.
For example, in the 2-point sampling method we
propagate the additional states $\ket{\psi^{\pm}}$ where
$\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$, with
$\sigma_{\lambda}$ being the standard deviation of $\lambda$.
The difference between the final states and the target
state is penalized, resulting in quadratic costs at
each knot point
${\lvert \ket{\psi^{\pm}_{k}} - \ket{\psi_{f}} \rvert}^{2}$.
For the 4-point sampling method we sample at two standard deviations
above and below the nominal parameter value in addition to one standard
deviation above and below.

(Comparison) To demonstrate the applicability of these techniques
to mitigate system parameter deviations,
we consider the task of achieving a single $X/2$
gate subject to a constant qubit frequency detuning
($f_{q} \gets f_{q} + \delta f_{q}$).
%% We use a compensating for off-resonance error
%% (CORPSE) dynamic decoupling pulse that is optimized
%% to mitigate first-order error arising due to the detuning.
%% The pulse is composed of
%% three rotations about the $\hat{x}$, $-\hat{x}$, and $\hat{x}$ axes
%% of the Bloch sphere
%% designed to compensate for erroneous rotations
%% (see Appendix).
We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
the sampling robustness methods accordingly.
For each method we compute the gate error for one simulated gate application
subject to the stated deviant dynamics.

(Results) Both the 2-point and 4-point
sampling methods converge on a similar solution with a low amplitude and
a single amplitude parity switch (see Figure 2a). At the gate time of the analytic pulse
($56.8$ ns) and a one percent qubit frequency detuning they achieve
gate errors $\sim 1.1 \cdot 10^{-4}$ commensurate with the analytic
pulse $\sim 1.2 \cdot 10^{-4}$. The 2-point and 4-point sampling methods
achieve similar metrics in simulation (see Table 2).

Both the 2\textsuperscript{nd}-order and 3\textsuperscript{rd}-order derivative methods
employ multiple amplitude parity switches and idle periods at the maximum
amplitude $0.5$ GHz. At the gate time of the analytic pulse and a one percent qubit
frequency detuning, the 2\textsuperscript{nd}-order method achieves a gate error
$\sim 0.8 \cdot 10^{-4}$ and the 3\textsuperscript{rd}-order method achieves a gate error
$\sim 1.1 \cdot 10^{-4}$ (see Figure 2b). Inspection of the simulation metrics reveals
that the 3\textsuperscript{rd}-order method is able to achieve a lower norm of the
3\textsuperscript{rd}-order state derivative at the cost of higher 1\textsuperscript{st}-order
and 2\textsuperscript{nd}-order derivatives than the 2\textsuperscript{nd}-order method (see Table 3).

Further, we observe that the derivative methods are able
to take advantage of an increased gate duration to achieve greater robustness,
while the sampling methods are unable to. The tradeoff between gate error and
gate duration appears roughly linear, where trading 3 times the gate duration
results in a factor of 4 reduction in gate error for the 2\textsuperscript{nd}-order
derivative method (using the 50ns and 150ns points, include this remark?). Additionally,
no optimization converged for gate durations less than 50ns., and robustness commensurate
with the analytic pulse at a shorter gate duration was not able to be achieved.

%% T2
\begin{table}
  \begin{tabular}{c | c | c}
    Method & ${\lvert \partial_{f_{q}} \ket{\psi_{N}} \rvert}^{2}$ ($10^{3}$)
    & ${\lvert \partial^{2}_{f_{q}} \ket{\psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
    \hline
    D-1 & 2.509 & 13.292\\
    D-2 & 3.711 & 1.452\\
  \end{tabular}
  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
    for $X/2$ gates optimized using the derivative methods. The norms are computed
    at the end of the gate's duration $t_{N} = 110$ns.}
\end{table}

\section{Robustness to Stochastic Parameter Deviations} %% S6
(Comparison) An additional task of interest is to mitigate
stochastic parameter deviations. For the fluxonium qubit, magnetic flux noise
acts to modify the flux amplitude from its intended value by an amount $\delta a$.
It is well studied that the spectral density of $\delta a$ follows a
1/$f$ distribution, consisting primarily of low frequency noise. We compare
the methods discussed for static parameter deviations to the task of
realizing a $X/2$ gate subject to 1/$f$ flux noise. We simulate successive
applications of the gate constructed by each method and measure the gate error
after each application (see Figure 3b). The system is subject to flux noise
$\delta a (t)$ that is generated from a standard normal distribution, normalized
by 1/$f^{1/2}$ in frequency space, and modulated by the experimentally obtained
flux noise amplitude $A_{\Phi} = 5.21 \mu \Phi_{0}$.

@@

@@

%% AE
\section{Dynamic Decoupling Sequences}
Here we present the equations for constructing the CORPSE and
B2CORPSE sequences.

%% AF
\section{Complex Tensor Handling}
We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
because the software we use does not support complex numbers yet.

>>>>

%% ABSTRACT
\begin{abstract}
  %% - 1 + 2 topic importance
  %% - 1 what I have done
  %% - few sentences about primary results
  The ability to engineer high fidelity operations on quantum processors in the presence of
  systematic errors and decoherence remains the primary challenge requisite to achieving quantum advantage.
  Quantum optimal control (QOC) techniques have proven effective in realizing high fidelity operations,
  but they require exquisite calibration to be performant.
  In this work we employ robust trajectory optimization techniques
  to achieve high fidelity gates for parameter spreads on the order of $1\%$. We propose a method that
  takes advantage of deviant parameter derivative information while maintaining
  computational efficiency by transforming
  higher-order differential equations into a set of coupled first-order equations.
  Additionally, current QOC techniques mitigate decoherence effects due to longitudinal relaxation by integrating
  the Lindblad master equation, which increases the computational complexity
  of the optimization algorithm.
  In this work we achieve a factor of $5$ increase in longitudinal relaxation times
  for the gate set of a fluxonium qubit
  using a computationally efficient metric in optimization.
\end{abstract}

@@

%% S2
\section{QOC + AL-iLQR}
(QOC Problem Statement) Here we introduce the notation
we will use throughout the paper,
review the quantum optimal control problem statement,
and introduce the trajectory optimization framework.
Quantum optimal control concerns the evolution of
a quantum state $\ket{\psi(t)}$ governed by the time-dependent
Schroedinger equation (TDSE)
\label{eq:tdse}
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H(u(t), t) \ket{\psi}
\end{align}
The evolution is sometimes cast with the evolution
of a density matrix under the Lindblad master equation to
model the decoherence of the state explicitly. The Hamiltonian
has an arbitrary dependence on the possibly multi-valued controls $u(t)$.
The controls are so called because they are the means the experimentalist has to
act on the system.

@@

(AL-iLQR Problem Statement) The trajectory optimization
literature solves a more general class of non-linear programs that resemble
the quantum optimal control problem. The quantum optimal control
problem is a specific case of the linear quadratic regulator (LQR).
LQR is so called because the dynamics are linear in the state and
the functional is quadratic in the state. In the LQR formulation
the same functional is evaluated at each knot point
\begin{equation}
  J_{\textrm{iLQR}} = \tilde{x}_{N}^{T} Q_{N} \tilde{x}_{N}
  + \sum_{k = 0}^{N - 1} \tilde{x}_{k}^{T} Q_{k} \tilde{x}_{k} + u_{k}^{T} R_{k} u_{k}
\end{equation}
where $\tilde{x}_{k} = x_{k} - x_{f}$ is the difference between the state
at knot point $k$ and final state, $u_{k}$ are the controls,
and $Q_{k}, R_{k}$ are matrices that define the penalty metric.
The state is propagated using a dynamics function
$x_{k + 1} = f(x_{k}, u_{k}, t_{k}, \Delta t_{k})$.
In the case of quantum optimal control $\ket{\psi_{k}} \subseteq x_{k}$
and $f$ encodes the TDSE dynamics. In the following
we refer to $\ket{\psi}$ as the state, $x$ as the augmented
state, and $u$ as the augmented controls.

@@

%% S4
\section{Longitudinal Relaxation Awareness}
(Strategy) We seek to minimize the probability
that the qubit decays as a result of longitudinal
relaxation. To this end we add the
longitudinal relaxation probability
to the augmented state vector
\begin{equation}
  P_{1}(t) = \int_{0}^{t} T_{1}^{-1}(a(t^{\prime})) dt^{\prime}
\end{equation}
Setting the target
probability to zero results in
a quadratic cost at each knot point
${\lvert P_{1}(t_{k}) \rvert}^{2}$.
$T_{1}(a_{k})$ is obtained at each knot point by evaluating
a spline fit to
experimental data of the form
$\{(a, T_{1})\}$.
Calculating $T_{1}$ directly from theoretical
considerations requires many high-rank
eigendecompositions, which
is computationally expensive. Additionally,
$T_{1}$ values are known to fluctuate greatly
with laboratory temperatures \cite{klimov2018fluctuations}.
Interpolating $T_{1}$ from experimental data increases
the fridge truth of the simulation.

For a constant $T_{1}$ time, a shorter gate duration would
favor a lower longitudinal relaxation probability.
We allow the optimizer to tune the gate duration by
making the time step between each knot point $\Delta t_{k}$
a decision variable.
Promoting $\Delta t_{k}$ to a decision variable, rather
than the number of knot points $N$, preserves the
Markovian decision structure of the trajectory
optimization problem. To ensure numerical
integration accuracy is maintained we add a bound
constraint at each knot point
$5\textrm{e-}2 \ \textrm{ns} \le
\Delta t_{k} \le 2\textrm{e-}1 \ \textrm{ns}$.
This bound constraint may be
violated for intermediate iterations of the optimization,
so we add the square root of the time step $\sqrt{\Delta t_{k}}$
to the augmented control vector and use the squared root
of the time step $\lvert \Delta t_{k} \rvert$ in optimization.

The optimizations for the $Y/2$ and $Z/2$ gates converge on similar solutions, a periodic
waveform with amplitude $\sim 0.2 \textrm{GHz}$, see Figure 1.
Both solutions extend their gate times
beyond their analytic counterpart, trading longer gate times for access
to higher amplitudes and therefore higher $T_{1}$ times. Both gates reduce
their probability of longitudinal relaxation per gate by a factor of $\sim 5$ over
their analytic counterparts, a similar scaling for their single gate error improvements
$\sim 3 \cdot 10^{-5} \rightarrow \sim 7 \cdot 10^{-6}$.
We define gate error as the infidelity of the final state and the target state
averaged over multiple pseudo-randomly generated initial states.
Furthermore, they perform simlarly in
the concatenated gate application comparision, maintaining a total gate error under $10^{-2}$
over $10^{3}$ gate applications. 
The $T_{1}$ optimized $X/2$ gate achieves a single gate error...

@@

  \caption{
    (a) Z/2 gates constructed using the
    analytic, 8-point sampling, 10-point unscented sampling, and the 1\textsuperscript{st}
    and 2\textsuperscript{nd} order derivative methods. The gates
    shown for the robust methods are the solutions at twice the analytic
    gate time.
    (b) Single gate error as
    a function of the gate duration at a one-percent
    detuning from the nominal qubit frequency for all methods.
    (c) Single gate error as a function of the detuning from the nominal
    qubit frequency at multiples of the analytic gate duration for the analytic
    and 1\textsuperscript{st} order derivative methods.
  }
\end{figure*}

@@

%% S5
\section{Robustness to Static Parameter Deviations}
Gate errors may be incurred if Hamiltonian parameters deviate
from their nominal values in the experiment. For example, the qubit frequency
may be measurable up to a known precision $f_{q} = \bar{f_{q}} \pm \sigma_{f_{q}}$
and drift in laboratory temperatures may increase
amplitude gain in the electronics chain $a \gets a (1 + \epsilon)$.
Multiple analytic methods have been presented to construct gates
that mitigate the error arising from parameter deviations
including composite pulse sequences \cite{merrill2014progress},
the DRAG scheme \cite{krantz2019quantum}, and
geometric phase considerations
\cite{xu2020nonadiabatic, han2020experimental}.
Composite pulse sequences are derived by computing
the erroroneous rotation arising from the static parameter deviation
with a finite order magnus expansion.
Pulses are then suitably composed to eliminate the
error. In principle the error may be eliminated
to arbitrary order with sufficiently many pulses.
It is difficult to choose an appropriate composite pulse
for comparison on the problem we study here. We
propose comparing numerical methods to existing analytic
techniques for future work.

We consider two numerical strategies for achieving robustness
to parameter deviations. The first we call the
derivative method. We draw on the intuition that
the sensitivity of the state evolution to a parameter
$\lambda$ is encoded in the $l$\textsuperscript{th}-order
derivative of the state with respect to that parameter
$\ket{\partial_{\lambda}^{l} \psi}$. The $m$\textsuperscript{th}-order
derivative method minimizes the norm of the first $m$
state derivatives with a quadratic cost at each knot point
${\lvert \braket{\partial^{l}_{\lambda} \psi_{k} | \partial^{l}_{\lambda}
    \psi_{k}} \rvert}^{2}$, $l \in \{1, \dots, m\}$.

The state derivatives could be obtained with backwards mode differentiation.
Naive automatic differentiation would compute
the state derivative at all $1, \dots, k - 1$ knot points
to obtain the state derivative at knot point $k$.
For $d$ sensitivity parameters and $N$ knot points
this requires $O(d N^2)$ matrix multiplications. Instead,
we forward propagate the state derivatives in the
augmented state vector under coupled equations of motion, resulting
in $O(d N)$ matrix multiplications.
For example, the equations of motion for the 1\textsuperscript{st}-order derivative
method are
\begin{align}
  i \hbar \frac{d}{dt} \ket{\psi} &= H \ket{\psi}\\
  i \hbar \frac{d}{dt} \ket{\partial_{\lambda}\psi} &=
  H \ket{\partial_{\lambda} \psi} +
  (\partial_{\lambda} H) \ket{\psi}
\end{align}
Matrix exponential schemes that account for the non-linear
term may be used to efficiently integrate the coupled dynamics
\cite{berland2005solving, einkemmer2017performance}.

The second scheme we analyze is the sampling method,
which has been proposed previously for
QOC \cite{reinhold2019controlling}.
We propagate additional states in the augmented state vector that follow deviant dynamics
and penalize their differences from the target state.
The states evolve under the deviant
Hamiltonian where the parameter of interest is replaced by
a deviant value $\lambda \gets \lambda + \delta \lambda$.
We propagate the additional states $\ket{\psi^{\pm}}$ with
deviant parameters $\lambda^{\pm} = \lambda \pm \sigma_{\lambda}$.
The differences between the final states and the target
state are penalized, resulting in quadratic costs at
each knot point ${\lvert \braket{\tilde{\psi}^{\pm}_{k} | \tilde{\psi}^{\pm}_{k}}
  \rvert}^{2}, \ket{\tilde{\psi}^{\pm}_{k}} = \ket{\psi^{\pm}_{k}} - \ket{\psi_{f}}$.
We sample two copies ($\ket{\psi^{\pm}}$) of four initial states with this method,
resulting in eight additional states added to the augmented state vector.
The four initial states are chosen so
that their outer products span the operators on the
Hilbert space $\{\ket{0}, \ket{1}, (\ket{0} + i\ket{1}) / \sqrt{2},
  (\ket{0} - \ket{1}) / \sqrt{2}\}$ \cite{chow2009randomized}. We observe
that sampling only the basis states $\{\ket{0}$, $\ket{1}\}$
is insufficient because, as in the case of a $Z$ rotation, their
trajectories are insensitive to some parameter deviations and therefore
they are undesirable samples for those operations.

Additionaly, we consider the unscented sampling method,
which is well-studied in the robust control iterature
\cite{julier2004unscented, lee2013sigma, manchester2016derivative}.
The strategy is to
propagate an ensemble of states (sigma points) which represent a distribution
over every element of $\ket{\psi}$. The distribution
models the uncertainty in each state element arising from the
parameter deviation. The ensemble consists of $2 (n + d)$ states
where $n$ is the size of the state vector and $d$ is the number of
deviant parameters. Each state in the ensemble is propagated to the next
knot point using separate deviant dynamics. Then, the mean and covariance of
the propagated ensemble is calculated. New states are sampled from the distribution
given by the calculated statistics for propagation to the next knot point.
This resampling procedure, the unscented transformation, accurately propagates
first and second moments of the distribution and ensures the points
lie on the $\sqrt{n}$\textsuperscript{th}
covariance contour at each knot point. A detailed update procedure is given
in the appendix. The initial state of all $2 (n + d)$ samples is
$(\ket{0} + i\ket{1} / \sqrt{2})$ with added noise. Repeating this procedure for the other
three operator basis states would provide the optimizer with more information,
but come at a computational cost which far exceeds that of the other methods,
resulting in $O(n^2)$ additional states in the augmented state vector.

To demonstrate the applicability of these techniques
to mitigate system parameter deviations,
we consider the task of achieving a single $Z/2$
gate subject to a constant qubit frequency detuning
($f_{q} \gets f_{q} + \delta f_{q}$).
We take $\sigma_{f_{q}} / f_{q} = 1\%$ to be one standard devation, and equip
the sampling methods accordingly. For each method we compute the gate error for
one simulated gate application subject to the deviant dynamics given by the
stated qubit frequency detuning.

We compare the robust methods to an analytically derived $Z/2$ pulse which corresponds to
idling at the flux frustration point $a = 0$. The analytic pulse
is at the device's speed limit for a $Z/2$ gate $t_{Z/2} = 1 / 4 f_{q}$ and
is simple to derive. It is linearly sensitive to qubit frequency detunings,
the erroneous rotation given by $2 \pi t_{Z/2} \delta f_{q}$. At a one-percent
qubit frequency detuning, it achieves a gate error of $\sim 4.5 \cdot 10^{-5}$, sufficient
for quantum error correction.

Each numerical method is able to find solutions at
all gate times above $t_{Z/2}$, but is unable to find solutions at shorter times.
The ability to perform $Z$ rotations in arbitrary times
is critical for operating multi-qubit experiments in the lab frame.
These numerical methods give an effective mechanism for synchronizing
qubits operating at different frequencies $f_{q, i} \neq f_{q, j}$.
The sampling and unscented sampling methods
converge on qualitatively similar solutions which combine idling periods
with fast ramps to the maximum amplitude. The gate error at a one-percent
detuning from the nominal qubit frequency achieved
by the sampling method does not improve substantially over the
range of gate durations. The unscented sampling method
achieves linear decreases in its gate error with longer gate durations
until half the larmor period $\sim 1 / 2 f_{q}$ after which it achieves a consistent
gate error $\sim 3.5 \cdot 10^{-5}$.

The derivative methods also converge on qualitatively similar solutions that
use fast traingle pulses at the boundaries and balance time
on either side of the flux-frustration point symmetrically at low amplitudes.
They achieve super-linear scaling in their gate error as
a function of the gate duration which approaches zero at four multiples
of $t_{Z/2}$. We observe that on a range of problems at small parameter detunings,
the 1\textsuperscript{st}-order derivative method tends to outperform the 2\textsuperscript{nd}-order
derivative method. We believe this is due to the low contribution of second-order
terms to the gate error in this deviation regime, see Table 2.

%% T2
%% TODO: Appendix
\begin{table}
  \begin{tabular}{c | c | c}
    Method & ${\lvert \ket{\partial_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{3}$)
    & ${\lvert \ket{\partial^{2}_{f_{q}} \psi_{N}} \rvert}^{2}$ ($10^{7}$)\\
    \hline
    D-1 & 2.051 & 17.935\\
    D-2 & 2.511 & 7.628\\
  \end{tabular}
  \caption{Norm of the first basis state's derivate with respect to the qubit frequency
    for $Z/2$ gates optimized using the derivative methods. The norms are computed
    at the end of the gate's duration $t_{N} = 38$ns.}
\end{table}

%% S6
\section{Robustness to Stochastic Parameter Deviations}
An additional source of error can be modeled by stochastic
Hamiltonian parameter deviations. For the fluxonium qubit, magnetic flux noise
modifies the flux amplitude from its nominal value by an amount $\delta a$.

It is well studied that the spectral density of $\delta a$ follows a
1/$f$ distribution, consisting primarily of low frequency noise. We compare
the methods discussed for static parameter deviations to the task of
realizing a $X/2$ gate subject to 1/$f$ flux noise. We simulate successive
applications of the gate constructed by each method and measure the gate error
after each application (see Figure 3b). The system is subject to flux noise
$\delta a (t)$ that is generated from a standard normal distribution, normalized
by 1/$f^{1/2}$ in frequency space, and modulated by the experimentally obtained
flux noise amplitude $A_{\Phi} = 5.21 \mu \Phi_{0}$.

@@

%% T1
\begin{table}[ht]
  \begin{tabular}{c | c | c | c | c | c | c}
         & Analytic & QOC & & Analytic & QOC & \\
    Gate & $P_{1}\ (10^{-5})$ & $P_{1}\ (10^{-5})$ & $P_{1\textrm{A}} / P_{1\textrm{Q}}$
    & GE $(10^{-5})$ & GE $(10^{-5})$ & GE$_{\textrm{A}}$ / GE$_{\textrm{Q}}$\\
    \hline
    Z/2 & 5.745  & 1.149 & 5.000 & 3.568  & 0.727 & 4.905\\
    Y/2 & 5.253  & 1.157 & 4.540 & 3.107  & 0.763 & 4.072\\
    X/2 & 16.251 & 4.488 & 3.621 & 10.542 & 3.563 & 3.449\\
  \end{tabular}
  \caption{Probability of longitudinal relaxation for each gate
    evaluated at the gate's duration.}
\end{table}

@@

%% AE
\section{Complex Tensor Handling}
We use an isomorphism $\mathcal{H}(\mathbb{C}^{n}) \cong \mathcal{H}(\mathbb{R}^{2n})$
because the software we use does not support complex numbers yet.
